---
title: "CaseStudy2"
author: "jjsmu"
date: "2023-12-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, echo=FALSE}

# Print Session Information 
sessionInfo()

library(aws.s3)
library(class)
library(caret)
library(e1071)
library(ggplot2)
library(GGally)
library(plotly)
library(tidyverse)
library(naniar)
library(ggthemes)
```

## Import the Case Study 2 data set

```{r import}

# Get the current working directory
getwd()

# List files in root dir
list.files()

# Fetch data set from local folder
df_cs2 = read.csv("data/CaseStudy2-data.csv", header = TRUE) 
df_cs2_no_attrition = read.csv("data/CaseStudy2CompSet No Attrition.csv", header = TRUE) 
df_cs2_no_salary = read.csv("data/CaseStudy2CompSet No Salary.csv", header = TRUE) 

```
## Tidy the Case Study 2 Data Set
```{r tidy, echo=FALSE}
# View data frames in a table
View(df_cs2)

# Print data frame metadata  
str(df_cs2)

# Print summary of the data frames
summary(df_cs2)

# Print the number of rows and columns
dim(df_cs2)

# Variables # missing
gg_miss_var(df_cs2)

# Returns vector of variables # missing
sapply(df_cs2, function(x) sum(is.na(x)))
```

## Transform the Case Study 2 Data Set
```{r transform, echo = FALSE}

# Simple Data frame with four variables
df <- df_cs2 %>% 
  select(ID, JobRole, JobSatisfaction, MonthlyIncome) 

# Data frame employee job information
df_job <- df_cs2 %>%
  #filter(Attrition == "Yes") %>%
  select(ID, Attrition, JobInvolvement, JobLevel, JobRole, JobSatisfaction)

# Data frame employee pay information
df_pay <- df_cs2 %>%
  #filter(Attrition == "Yes") %>%
  select(ID, DailyRate, HourlyRate, MonthlyRate, MonthlyIncome, OverTime, PercentSalaryHike, StockOptionLevel, StandardHours)

# Data frame employee basic information
df_emp <- df_cs2 %>% 
  #filter(Attrition == "Yes") %>%
  select(ID, Age, Gender, Education, EducationField, Department, DistanceFromHome, MaritalStatus, NumCompaniesWorked, PerformanceRating, TotalWorkingYears, WorkLifeBalance, YearsAtCompany, YearsInCurrentRole, YearsWithCurrManager, YearsSinceLastPromotion)
```

## Visualize the Case Study 2 Data Set
### Perform Exploratory Data Analysis
```{r eda}
# Count the number of employees with attrition, we will compare with rest of sample later
sum(df_cs2$Attrition == "Yes") 

### Company Profile Snapshot: Age, Gender, Education, Pay, Role ###

# Box plot age by department
box_plot_age_dept <- df_emp %>% ggplot(aes(Department, Age, fill = Department)) + geom_boxplot() + ggtitle(paste("Box plot of Age by Department, n = ", count(df_emp)))
ggplotly(box_plot_age_dept)

# Box plot years in current role by department
box_plot_yirc_dept <- df_emp %>% ggplot(aes(Department, YearsInCurrentRole, fill = Department)) + geom_boxplot() + ggtitle(paste("Box plot of Years In Current Role by Department, n = ", count(df_emp)))
ggplotly(box_plot_yirc_dept)

# Box plot age by job role
box_plot_age_role <- merge(df_emp, df_job) %>% ggplot(aes(JobRole, Age, fill = JobRole)) + geom_boxplot() + ggtitle(paste("Box plot of Age by Job Role, n = ", count(df_emp))) + theme(axis.text.x = element_text(angle = 15, vjust = 0.5, hjust = 1))
ggplotly(box_plot_age_role)

# Box plot years in current role by job role
box_plot_role_yicr <- merge(df_emp, df_job) %>% ggplot(aes(JobRole, YearsInCurrentRole, fill = JobRole)) + geom_boxplot() + ggtitle(paste("Box plot of Years in current role by Job Role, n = ", count(df_emp))) + xlab("Job Role") + ylab("Years in current role") + theme(axis.text.x = element_text(angle = 15, vjust = 0.5, hjust = 0.5)) 
ggplotly(box_plot_role_yicr) 

# Bar plot job role by gender
bar_plot_dodge_role_gender <- merge(df_job, df_emp) %>% ggplot(aes(JobRole, fill = Gender)) + geom_bar(position = "dodge") + ggtitle(paste("Bar plot of Job Roles by Gender,  n = ", count(df))) + xlab("Job Role") + ylab("Count") + theme(axis.text.x = element_text(angle = 15, vjust = 0.5, hjust = 0.5))
ggplotly(bar_plot_dodge_role_gender)

# Box plot of monthly income by job role
box_plot_income_role <- df %>% ggplot(aes(JobRole, MonthlyIncome, fill = JobRole)) + geom_boxplot() + ggtitle(paste("Box plot of Monthly Income by Job Role, n = ", count(df))) + xlab("Job Role") + ylab("Monthly Income (USD)") + theme(axis.text.x = element_text(angle = 15, vjust = 0.5, hjust = 0.5)) 
ggplotly(box_plot_income_role) 

# Bar plot of job satisfaction by job role
bar_plot_dodge_job_sat <- df %>% ggplot(aes(JobSatisfaction, fill = JobRole)) + geom_bar(position = "dodge") + ggtitle(paste("Bar plot of Job Satisfaction by Job Roles,  n = ", count(df))) + xlab("Job Satisfaction") + ylab("Count") + theme(axis.text.x = element_text(angle = 15, vjust = 0.5, hjust = 1))
ggplotly(bar_plot_dodge_job_sat)

# Histogram of monthly income
hist_wrap_income_role <- df %>% ggplot(aes(MonthlyIncome, fill = JobRole)) + geom_histogram(binwidth = 200) + ggtitle(paste("Histogram of Monthly Income by Job Role, n = ", count(df))) + xlab("Monthly Income (USD)") + ylab("Count") + theme(axis.text.x = element_text(angle = 15, vjust = 0, hjust = 0.1)) + facet_wrap(~JobRole)
ggplotly(hist_wrap_income_role)

# Histogram of job satisfaction by job role
hist_wrap_income_sat <- df %>% ggplot(aes(MonthlyIncome, fill = JobRole, position = "stack")) + geom_histogram(binwidth = 200) + ggtitle(paste("Histogram of Monthly Income by Job Satisfaction, n = ", count(df))) + xlab("Monthly Income (USD)") + ylab("Count") + facet_wrap(~JobSatisfaction)
ggplotly(hist_wrap_income_sat)

### Attritional Factors by Variables ###

# Bar plot of attrition by job role
hist_att_role <- merge(df_emp, df_job) %>% ggplot(aes(JobRole, fill = Attrition)) + geom_bar(position = "stack") + ggtitle(paste("Bar plot of Attrition by Job Role, n = ", count(df_emp))) + xlab("Job Role") + ylab("Count") + theme(axis.text.x = element_text(angle = 15, vjust = 0, hjust = 1))
ggplotly(hist_att_role)

# Bar plot of attrition v. gender by job role
hist_wrap_role_gender <- merge(df_emp, df_job) %>% ggplot(aes(Attrition, fill = Gender)) + geom_bar(position = "dodge") + ggtitle(paste("Bar plot of Attrition by Gender, n = ", count(df_emp))) + xlab("Attrition") + ylab("Count") + facet_wrap(~JobRole)
ggplotly(hist_wrap_role_gender)

# Plot of performance v. distance from home by job role
smooth_wrap_dist_rating <- merge(df_emp, df_job) %>% ggplot(aes(DistanceFromHome, PerformanceRating, color = JobRole)) + geom_smooth() + facet_wrap(~JobRole)
ggplotly(smooth_wrap_dist_rating)

# Plot of performance v. number of companies worked by job role
smooth_wrap_numcos_rating <- merge(df_emp, df_job) %>% ggplot(aes(NumCompaniesWorked, PerformanceRating, color = JobRole)) + geom_smooth() + facet_wrap(~JobRole)
ggplotly(smooth_wrap_numcos_rating)

# Plot of performance v. years since last promotion by job role
smooth_wrap_promo_rating <- merge(df_emp, df_job) %>% ggplot(aes(YearsSinceLastPromotion, PerformanceRating, color = JobRole)) + geom_smooth() + facet_wrap(~JobRole)
ggplotly(smooth_wrap_promo_rating)

# PercentHike v. Years In Current Role
plot_wrap_psh_yslp <- df_cs2 %>% ggplot(aes(YearsSinceLastPromotion, PercentSalaryHike)) + geom_point(aes(color = Attrition), size = 0.2, position = "jitter") + facet_wrap(~JobRole)
ggplotly(plot_wrap_psh_yslp)

# Scatter plot years since last promotion v. years in current role
plot_wrap_yslp_yicr <- df_cs2 %>% ggplot(aes(YearsSinceLastPromotion, YearsInCurrentRole)) + geom_point(aes(color = Attrition), size = 0.2, position = "jitter") + facet_wrap(~JobRole)
ggplotly(plot_wrap_yslp_yicr)

# *** Scatter plot years with current manager v. years in current role ***
plot_wrap_ywcm_yicr <- df_cs2 %>% ggplot(aes(YearsWithCurrManager, YearsInCurrentRole)) + geom_point(aes(color = Attrition), size = 0.2, position = "jitter") + facet_wrap(~JobRole)
ggplotly(plot_wrap_ywcm_yicr)

# Scatter plot years with current manager v. job satisfaction
plot_wrap_ywcm_js <- df_cs2 %>% ggplot(aes(YearsWithCurrManager, JobSatisfaction)) + geom_point(aes(color = Attrition), size = 0.2, position = "jitter") + facet_wrap(~JobRole)
ggplotly(plot_wrap_ywcm_js)
```

## Model the Case Study 2 data frame
### KNN Classification
```{r knn}
#source("analysis/knn.R")

# Classification - attrition by leadership management
set.seed(6)
split_percent <- 0.7

l_train <- df_cs2 %>% 
  select(Attrition, YearsWithCurrManager, YearsInCurrentRole)

trainIndices = sample(1:dim(l_train)[1], round(split_percent * dim(l_train)[1]))
train = l_train[trainIndices,]
test = l_train[-trainIndices,]

scatter_smooth_ywcm_yicr <- l_train %>% ggplot(aes(YearsWithCurrManager, YearsInCurrentRole, color = Attrition)) + geom_point(position = "jitter") + geom_smooth(aes(color = Attrition)) + ggtitle(paste("Plot of Years with current manager v. Years in current role, n = ", count(l_train)))
ggplotly(scatter_smooth_ywcm_yicr)

classification <- knn(train[,2:3], test[,2:3], train$Attrition, prob = TRUE, k = 5)
table(classification, test$Attrition)
confusionMatrix(table(classification, test$Attrition))

# Classification - attrition by time duration
set.seed(6)
split_percent <- 0.7

l_train <- df_cs2 %>% 
  select(Attrition, YearsAtCompany, YearsInCurrentRole)

trainIndices = sample(1:dim(l_train)[1], round(split_percent * dim(l_train)[1]))
train = l_train[trainIndices,]
test = l_train[-trainIndices,]

scatter_smooth_yac_yicr <- l_train %>% ggplot(aes(YearsAtCompany, YearsInCurrentRole, color = Attrition)) + geom_point(position = "jitter") + geom_smooth(aes(color = Attrition)) + ggtitle(paste("Plot of Years at company v. years in current role, n = ", count(l_train)))
ggplotly(scatter_smooth_yac_yicr)

classification <- knn(train[,2:3], test[,2:3], train$Attrition, prob = TRUE, k = 5)
table(classification, test$Attrition)
confusionMatrix(table(classification, test$Attrition))

# Classification - attrition by peer relationship
set.seed(6)
split_percent <- 0.7

l_train <- df_cs2 %>% 
  select(Attrition, RelationshipSatisfaction, YearsWithCurrManager)

trainIndices = sample(1:dim(l_train)[1], round(split_percent * dim(l_train)[1]))
train = l_train[trainIndices,]
test = l_train[-trainIndices,]

scatter_smooth_ywcm_yicr <- l_train %>% ggplot(aes(RelationshipSatisfaction, YearsWithCurrManager, color = Attrition)) + geom_point(position = "jitter") + geom_smooth(aes(color = Attrition)) + ggtitle(paste("Plot of Number of companies v. years at company, n = ", count(l_train)))
ggplotly(scatter_smooth_ywcm_yicr)

classification <- knn(train[,2:3], test[,2:3], train$Attrition, prob = TRUE, k = 5)
table(classification, test$Attrition)
confusionMatrix(table(classification, test$Attrition))

```
### Naive Bayes Classification
```{r naivebayes}
#source("analysis/nb.R")

# Continuous predictor of attrition by leadership
df_attr <- df_cs2 %>%
  select(ID, Attrition, YearsWithCurrManager, YearsInCurrentRole) %>%
  mutate(ID = as.factor(ID), YearsInCurrentRole = as.factor(YearsInCurrentRole), YearsWithCurrManager = as.factor(YearsWithCurrManager))

df_attr %>% ggplot(aes(YearsWithCurrManager, YearsInCurrentRole)) + geom_point(aes(color = Attrition), position = "jitter", size = 0.3) + geom_smooth(aes(color = Attrition), size = 0.3)

iterations = 100
m_acc <- matrix(nrow = iterations)
split_percent <- 0.7
for(i in 1:iterations) {
  train_indices = sample(1:dim(df_attr)[1], round(split_percent * dim(df_attr)[1]))
  train = df_attr[train_indices,]
  test = df_attr[-train_indices,]
  
  model = naiveBayes(train[,3:4], train$Attrition)
  table(predict(model, test[,3:4]), test$Attrition)
  cm <- confusionMatrix(table(predict(model,test[,3:4]), test$Attrition))
  m_acc[i] <- cm$overall[1]
}

mean_acc <- colMeans(m_acc)
mean_acc

```

### Linear Regression Analysis
```{r regression}
#source("analysis/lm.R")

# Linear Model Monthly Income v. Monthly Rate
fit <- lm(MonthlyIncome~HourlyRate, data = df_pay)

df_pay %>% ggplot(aes(MonthlyIncome, MonthlyRate)) + geom_point() + geom_smooth(method = "lm")

beta_0_hat <- fit$coefficients[1]
beta_1_hat <- fit$coefficients[2]

SE_beta_0_hat <- summary(fit)$coefficients[1,2]
SE_beta_1_hat <- summary(fit)$coefficients[2,2]

# Intercept
tstat_int <- beta_0_hat / SE_beta_0_hat
pvalue_int <- (1-pt(tstat_int, length(df_pay$MonthlyIncome)-2)) * 2
tstat_int
pvalue_int

# Slope
tstat_slope <- beta_1_hat / SE_beta_1_hat
pvalue_slope <- (pt(tstat_slope, length(df_pay)-2)) * 2
tstat_slope
pvalue_slope

summary(fit)
confint(fit)

# Welch Two Sample t-test
t.test(df_pay$MonthlyIncome, df_pay$MonthlyRate)

# Conduct Hypothesis Test

# LOOCV
pred_error_sq <- c(0)
for(i in 1:dim(df_pay)[1]) {
  loocv_i <- df_pay[-i,]
  fit <- lm(MonthlyIncome ~ MonthlyRate, data = loocv_i)
  pred_i <- predict(fit, data.frame(MonthlyRate = df_pay[i,4]))
  pred_error_sq <- pred_error_sq + (df_pay[i,4] - pred_i)^2
}
SSE <- var(df_pay$MonthlyIncome)
R_squared <-  1 - (pred_error_sq/SSE)
MSE <- pred_error_sq / length(df_pay)
RMSE <- sqrt(pred_error_sq/length(df_pay))
RMSE

